---
title: "Appendix"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, echo = F, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tseries)
library(plm)
library(kableExtra)
```

\begin{center}
Practical Example - Assessing Random Walks and Autoregression Before Fitting CEMs
\end{center}


\newpage

The data were created via a true CEM process, consistent with prior work in Lang, Bliese, and Runge (2021). The data are stored in an object called `df`. Load the dataframe into your R session with

```{r}

df <- read.csv(
  "https://raw.githubusercontent.com/cdishop/spurious-consensus/master/data/cem.csv"
  )

```


\

Data Snippet:

\

```{r, echo = F}

df[1:5, ] %>%  kbl(booktabs = T) %>% kable_styling(latex_options = "hold_position")

```

## Step 1

In the first step, observations are checked against random walks.

\

Check the series of a single individual for a random walk:

```{r, warning = F, message = F}
library(tidyverse)
library(tseries)
library(plm)

oneperson <- df %>% 
  filter(id == unique(df$id)[1]) %>% 
  filter(group == unique(df$group)[1])

adf.test(oneperson$y)
```

\

Based on the test result, the trajectory of a single individual is inconsistent with a random walk. 

\newpage

Conduct the same test across all individuals and groups.

```{r, warning = F, message = F, eval = F}

people <- length(unique(df$id))
groups <- length(unique(df$group))
savetests <- numeric(people*groups)
counter <- 0
for(k in 1:groups){
  
  for(person in 1:people){
    counter <- counter + 1
    
    usedf <- df %>% 
      filter(id == unique(df$id)[person]) %>% 
      filter(group == k)
    
    result <- adf.test(usedf$y)$p.value
    savetests[counter] <- result
    
  }
  
}

sum(savetests < 0.05) / counter

```
```{r, echo = F}

print('[1] 0.87')

```

87% were not consistent with random walks. 

\newpage

## Step 2

In the second step, observations are checked for autoregressive effects.

```{r, warning = F, message = F, eval = T}
library(nlme)
noauto <- lme(y ~ 1,
              random = ~ time | group / id,
              data = df,
              control = list(opt = "optim"))
withauto <- lme(y ~ 1,
                random = ~ time | group / id,
                data = df,
                control = list(opt = "optim"),
                correlation = corARMA(p = 1))

anova(noauto, withauto)
```


\

The unconditional model without autoregressive effects is retained. 

## Step 3

Having assessed random walks and autoregressive effects, one can proceed with CEM. Fitting CEM models is well-described in existing articles. See in-text references. 

```{r, eval = F}

library(nlme)
step1<-lme(y ~ time, random = list(group=pdDiag(~1),id=pdSymm(~1)),
           data=df)
step2a<-update(step1,random = list(group=pdDiag(~time),id=pdSymm(~1))) 
step2b<-update(step2a,random = list(group=pdSymm(~time),id=pdSymm(~1))) 
step3<-update(step2b,weights=varExp(form = ~ time),
              control = list(opt = "optim"))
summary(step3)$modelStruct$varStruct 

```

